---
title: Mini Data Analysis 1
author: Jana Osea
date: October 5, 2021
output: github_document
---

# 0 Introduction

This file goes through my exploration of different data sets in the `datateachr` package. I will look at 7 data sets as described [here](https://stat545.stat.ubc.ca/mini-project/mini-project-1/). I will be documenting how I choose my final data set and then I will perform various data exploration techniques to this data set in order to create 4 research questions motivated by the exploration.

# 1 Choose Favorite Data set

## 1.1 Choose 4 Data sets

```{r, include = FALSE}
library(datateachr)
library(tidyverse)
library(dplyr)
```

From an initial pass through of the description of the various data set, I am most interested in the following 4 data sets:

1. *apt_buildings*
2. *building_permits*
3. *parking_meters*
4. *vancouver_trees*

## 1.2 Exploring Attributes of 4 Datasets

### 1.2.1 *apt_buildings*
```{r, echo=FALSE}
str(apt_buildings)
```

### 1.2.2 *building_permits*
```{r, echo=FALSE}
str(building_permits)
```
  
### 1.2.3 *parking_meters*
```{r, echo=FALSE}
str(parking_meters)
```

### 1.2.4 *vancouver_trees*
```{r, echo=FALSE}
str(vancouver_trees)
```
## 1.3 Choose 2 Data sets

## 1.4 Final Data set

# 2 Exploring the Data set


