---
title: "Mini Data Analysis Milestone 2"
output: github_document
---

*To complete this milestone, you can either edit [this `.rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-2.Rmd) directly. Fill in the sections that are commented out with `<!--- start your work here--->`. When you are done, make sure to knit to an `.md` file by changing the output in the YAML header to `github_document`, before submitting a tagged release on canvas.*

# Welcome back to your mini data analysis project!

This time, we will explore more in depth the concept of *tidy data*, and hopefully investigate further into your research questions that you defined in milestone 1.

**NOTE**: The main purpose of the mini data analysis is to integrate what you learn in class in an analysis. Although each milestone provides a framework for you to conduct your analysis, it's possible that you might find the instructions too rigid for your data set. If this is the case, you may deviate from the instructions -- just make sure you're demonstrating a wide range of tools and techniques taught in this class.

Begin by loading your data and the tidyverse package below: 

```{r, message = FALSE}
suppressPackageStartupMessages(library(datateachr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggridges))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(kableExtra))
```

# Learning Objectives

By the end of this milestone, you should:

+ Become familiar with manipulating and summarizing your data in tibbles using `dplyr` and `tidyr`, with a research question in mind.
+ Understand what *tidy* data is, and how to create it. In milestone 3, we will explore when this might be useful.
+ Generate a reproducible and clear report using R Markdown.
+ Gain a greater understanding of how to use R to answer research questions about your data.

**Things to keep in mind**

+ Remember to document your code, be explicit about what you are doing, and write notes in this markdown document when you feel that context is required. Create your analysis as if someone else will be reading it! **There will be 2.5 points reserved for reproducibility, readability, and repo organization.**

+ Before working on each task, you should always keep in mind the specific **research question** that you're trying to answer.

# Task 1: Process and summarize your data (15 points)

From milestone 1, you should have an idea of the basic structure of your dataset (e.g. number of rows and columns, class types, etc.). Here, we will start investigating your data more in-depth using various data manipulation functions. 

### 1.1 (2.5 points) 

First, write out the 4 research questions you defined in milestone 1 were. This will guide your work through milestone 2:

1. Do certain property types prefer building apartments with a certain number of units?
2. Is there are trend on the number of units built in smoking apartment buildings from 1910 to 2020?
3. Does the number of units change according to property type for certain visitor parking types?
4. For apartments with indoor pools, does the total number of units change according to ward?


### 1.2 (10 points)

Now, for each of your four research questions, choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing). You should have 2 tasks done for each research question (8 total). Make sure it makes sense to do them! (e.g. don't use a numerical variables for a task that needs a categorical variable.). Comment on why each task helps (or doesn't!) answer the corresponding research question. 

Ensure that the output of each operation is printed! 

**Summarizing:**

1. Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
2. Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
3. Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".*
4. Based on two categorical variables, calculate two summary statistics of your choosing.

**Graphing:**

5. Create a graph out of summarized variables that has at least two geom layers.
6. Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read. 
7. Make a graph where it makes sense to customize the alpha transparency.
8. Create 3 histograms out of summarized variables, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Make sure it's clear what research question you are doing each operation for!

#### 1.2.1 Research Question 1

The first research question is: Do certain property types prefer building apartments with a certain number of units?

##### 1.2.1.1 Task 1: Summary Statistics

The task is to compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data. This allows us to explore the distribution of values for number of units within different property types.

```{r}
# determine outliers
outliers <- boxplot.stats(apt_buildings$no_of_units)$out 

# make tibble data frame 
apt_buildings %>% 
  filter(!no_of_units %in% outliers) %>% # get rid of outliers
  group_by(property_type) %>%
  summarise(
    count = n(),
    mean = mean(no_of_units),
    sum = sum(no_of_units),
    ) -> apt_buildings1

# show the summary statistics
apt_buildings %>% 
  filter(!no_of_units %in% outliers) %>% # get rid of outliers
  group_by(property_type) %>%
  summarise(
    count = n(),
    range = range(no_of_units),
    mean = mean(no_of_units),
    sum = sum(no_of_units),
    )
```



##### 1.2.1.2 Task 5: Graph Summary Statistics

The task is to create a graph out of summarized variables that has at least two geom layers. I chose to use a violin plot as it shows the data distribution of the number of units over the different property types in an explicit way. In addition, we are able to explicitly see the different quantiles.

```{r}
apt_buildings %>% 
  filter(!no_of_units %in% outliers) %>% # get rid of outliers
  group_by(property_type) %>%
  ggplot(aes(x=property_type, y=no_of_units, color=property_type)) +
  geom_violin()+ 
  geom_boxplot(width=0.1)
```

#### 1.2.2 Research Question 2

The second research question is: Is there are trend on the number of units built in smoking apartment buildings from 1910 to 2020?

##### 1.2.2.1 Tasks 2: 

The task is to compute the number of observations for at least one of your categorical variables. Do not use the function `table()`! In this case, I created 3 different tibble data frames that count the data by different bins of years. `apt_buildings2.1` , `apt_buildings2.2`, and `apt_buildings2.3` are grouped by 10 years, 5 years, and 1 year correspondingly. The reason I did this was so that I can create visualizations as a secondary task to determine which bin would be the most useful ie. convey the same information without loss of generality. In addition, I added the extra column for 

```{r}
# task 3
apt_buildings %>%
  filter(!is.na(`non-smoking_building`)) %>%
  filter(!no_of_units %in% outliers) %>%
  mutate(year = as.Date(as.character(year_built), format="%Y")) %>%
  filter(year > as.Date("1910", format="%Y")) %>%
  mutate(year_category=cut(year_built, breaks=c(-Inf,seq(1910, 2020, 10)), labels=paste(c(seq(1900, 2010, 10)), c(seq(1909, 2009, 10), 2020), sep="-"))) %>%
  group_by(year_category, property_type) %>%
  count(`non-smoking_building`) -> apt_buildings2.1 # bin by 10

apt_buildings %>%
  filter(!is.na(`non-smoking_building`)) %>%
  filter(!no_of_units %in% outliers) %>%
  mutate(year = as.Date(as.character(year_built), format="%Y")) %>%
  filter(year > as.Date("1910", format="%Y")) %>%
  mutate(year_category=cut(year_built, breaks=c(-Inf, seq(1905, 2020, 5)), labels=paste(c(seq(1900, 2015, 5)), c(seq(1904, 2014, 5),2020), sep="-"))) %>%
  group_by(year_category, property_type) %>%
  count(`non-smoking_building`) -> apt_buildings2.2 # bin by 5


apt_buildings %>%
  filter(!is.na(`non-smoking_building`)) %>%
  filter(!no_of_units %in% outliers) %>%
  mutate(year = as.Date(as.character(year_built), format="%Y")) %>%
  filter(year > as.Date("1910", format="%Y")) %>%
  group_by(year_built, property_type) %>%
  count(`non-smoking_building`) -> apt_buildings2.3 # bin by 1

apt_buildings2.1
apt_buildings2.2
apt_buildings2.3
```



##### 1.2.2.2 

The task is to create 3 histograms out of summarized variables, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

```{r}
# task 8
apt_buildings2.1 %>%
  ggplot(aes(x=year_category, y=n, fill=`non-smoking_building`)) +  
  geom_bar(stat="identity")

apt_buildings2.2 %>%
  ggplot(aes(x=year_category, y=n, fill=`non-smoking_building`)) +  
  geom_bar(stat="identity")

apt_buildings2.3 %>%
  ggplot(aes(x=year_built, y=n, fill=`non-smoking_building`)) +  
  geom_bar(stat="identity")
```


#### 1.2.3 Research Question 3

The second research question is: Do certain property types prefer building apartments with a certain number of units?

##### 1.2.3.1

The task is to compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!

```{r}
# task 2
apt_buildings %>%
  filter(!is.na(visitor_parking)) %>%
  filter(!is.na(property_type)) %>%
  filter(!visitor_parking == "UNAVAILABLE") %>% 
  group_by(property_type, visitor_parking) %>%
  summarise(no_of_units_sum=sum(no_of_units)) %>%
  mutate(no_of_units_sum_log = log(no_of_units_sum))-> apt_buildings3
apt_buildings3
```

##### 1.2.3.2

The task is to create a graph out of summarized variables that has at least two geom layers.

```{r}
# task 5
apt_buildings3 %>% 
  ggplot(aes(x=property_type, y=no_of_units_sum, colour=visitor_parking, group=visitor_parking)) +
  geom_line() +
  geom_point()

apt_buildings3 %>% 
  ggplot(aes(x=property_type, y=no_of_units_sum_log, colour=visitor_parking, group=visitor_parking)) +
  geom_line() +
  geom_point()
```

#### 1.2.4 Research Question 4

##### 1.2.4.1

The task is to compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!

```{r}
# task 2
# count number of amenities by splitting words
no_of_amenities <- rep(0, nrow(apt_buildings))
for(i in 1:nrow(apt_buildings)) {
  amenities = apt_buildings$amenities[i]
  if(!is.na(amenities)) {
    no_of_amenities[i] = length(str_split(amenities, " ,")[[1]])
  }
}

apt_buildings %>%
  cbind(no_of_amenities) %>%
  na.omit() %>%
  group_by(ward, `non-smoking_building`) %>%
  summarise(no_of_amenities_sum=sum(no_of_amenities)) -> apt_buildings4
```

##### 1.2.4.2

The task is to create a graph out of summarized variables that has at least two geom layers.

```{r}
# task 5
apt_buildings4  %>%
  ggplot(aes(x=ward, y=no_of_amenities_sum, fill=`non-smoking_building`)) +
  geom_bar(stat="identity") +
  facet_wrap(~`non-smoking_building`)
```

### 1.3 (2.5 points)

Based on the operations that you've completed, how much closer are you to answering your research questions? Think about what aspects of your research questions remain unclear. Can your research questions be refined, now that you've investigated your data a bit more? Which research questions are yielding interesting results?

<!------------------------- Write your answer here ---------------------------->


<!----------------------------------------------------------------------------->

# Task 2: Tidy your data (12.5 points)

In this task, we will do several exercises to reshape our data. The goal here is to understand how to do this reshaping with the `tidyr` package.

A reminder of the definition of *tidy* data:

- Each row is an **observation**
- Each column is a **variable**
- Each cell is a **value**

*Tidy'ing* data is sometimes necessary because it can simplify computation. Other times it can be nice to organize data so that it can be easier to understand when read manually. 

### 2.1 (2.5 points)

Based on the definition above, can you identify if your data is tidy or untidy? Go through all your columns, or if you have >8 variables, just pick 8, and explain whether the data is untidy or tidy.

<!--------------------------- Start your work below --------------------------->


<!----------------------------------------------------------------------------->

### 2.2 (5 points)

Now, if your data is tidy, untidy it! Then, tidy it back to it's original state.

If your data is untidy, then tidy it! Then, untidy it back to it's original state.

Be sure to explain your reasoning for this task. Show us the "before" and "after".

<!--------------------------- Start your work below --------------------------->


<!----------------------------------------------------------------------------->

### 2.3 (5 points)

Now, you should be more familiar with your data, and also have made progress in answering your research questions. Based on your interest, and your analyses, pick 2 of the 4 research questions to continue your analysis in milestone 3, and explain your decision. 

Try to choose a version of your data that you think will be appropriate to answer these 2 questions in milestone 3. Use between 4 and 8 functions that we've covered so far (i.e. by filtering, cleaning, tidy'ing, dropping irrelvant columns, etc.). 

fdfasdfasdf

<!----------------------------------------------------------------------------->

*When you are done, knit an `md` file. This is what we will mark! Make sure to open it and check that everything has knitted correctly before submitting your tagged release.*

### Attribution

Thanks to Victor Yuan for mostly putting this together. 