---
title: "Mini Data Analysis 2"
author: Jana Osea
date: October 18, 2021
output: github_document
---

# 0 Introduction

This file goes through my data exploration to answer some questions as outlined in the [first mini data analysis](https://github.com/stat545ubc-2021/jana-mini-data-analysis/blob/main/mini-data-analysis-1.md). The data used is from the  `datateachr` package. The project outline can be [here](https://stat545.stat.ubc.ca/mini-project/mini-project-2/). I will be documenting the data exploration tasks used for each research question. First, we load the appropriate library packages.


```{r, message = FALSE}
suppressPackageStartupMessages(library(datateachr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggridges))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(kableExtra))
```

# 1 Process and summarize your data 

## 1.1 Research Questions 

First, I write out the 4 research questions defined in milestone 1. 

1. Do certain property types prefer building apartments with a certain number of units?
2. Is there are trend on the number of units built in smoking apartment buildings from 1910 to 2020?
3. Does the number of units change according to property type for certain visitor parking types?
4. For apartments with indoor pools, does the total number of units change according to ward?


## 1.2 Summarizing and Graphing Tasks

Now, for each of four research questions, I choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing).

**Summarizing:**

1. Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
2. Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
3. Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".*
4. Based on two categorical variables, calculate two summary statistics of your choosing.

**Graphing:**

5. Create a graph out of summarized variables that has at least two geom layers.
6. Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read. 
7. Make a graph where it makes sense to customize the alpha transparency.
8. Create 3 histograms out of summarized variables, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.


### 1.2.1 Research Question 1

The first research question is: Do certain property types prefer building apartments with a certain number of units?

#### 1.2.1.1 Task 1: Summary Statistics

The task is to compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data. This allows us to explore the distribution of values for number of units within different property types.

```{r}
# determine outliers
outliers <- boxplot.stats(apt_buildings$no_of_units)$out 

# make tibble data frame 
apt_buildings %>% 
  filter(!no_of_units %in% outliers) %>% # get rid of outliers
  group_by(property_type) %>%
  summarise(
    count = n(),
    mean = mean(no_of_units),
    sum = sum(no_of_units),
    ) -> apt_buildings1

# show the summary statistics
apt_buildings %>% 
  filter(!no_of_units %in% outliers) %>% # get rid of outliers
  group_by(property_type) %>%
  summarise(
    count = n(),
    range = range(no_of_units),
    mean = mean(no_of_units),
    sum = sum(no_of_units),
    )
```



#### 1.2.1.2 Task 5: Graph Summary Statistics

The task is to create a graph out of summarized variables that has at least two geom layers. I chose to use a violin plot as it shows the data distribution of the number of units over the different property types in an explicit way. In addition, we are able to explicitly see the different quantiles.

```{r}
apt_buildings %>% 
  filter(!no_of_units %in% outliers) %>% # get rid of outliers
  group_by(property_type) %>%
  ggplot(aes(x=property_type, y=no_of_units, color=property_type)) +
  geom_violin()+ 
  geom_boxplot(width=0.1)
```

### 1.2.2 Research Question 2

The second research question is: Is there are trend on the number of units built in smoking apartment buildings from 1910 to 2020?

#### 1.2.2.1 Tasks 2: Compute observations

The task is to compute the number of observations for at least one of your categorical variables. Do not use the function `table()`! In this case, I created 3 different tibble data frames that count the data by different bins of years. `apt_buildings2.1` , `apt_buildings2.2`, and `apt_buildings2.3` are grouped by 10 years, 5 years, and 1 year correspondingly. The reason I did this was so that I can create visualizations as a secondary task to determine which bin would be the most useful ie. convey the same information without loss of generality. It is also good to note that I am summing over the *no_of_units* of each group since that is a more accurate number to count rather than just the counts of each observation. In addition, I added the extra column of describing property type as I potentially would like to explore how this factor can affect the number smoking units.

```{r}
# task 3
apt_buildings %>%
  filter(!is.na(`non-smoking_building`)) %>%
  filter(!no_of_units %in% outliers) %>%
  mutate(year = as.Date(as.character(year_built), format="%Y")) %>%
  filter(year > as.Date("1910", format="%Y")) %>%
  mutate(year_category=cut(year_built, breaks=c(-Inf,seq(1910, 2020, 10)), labels=paste(seq(1900, 2010, 10)))) %>%
  group_by(year_category, property_type) %>%
  count(`non-smoking_building`) -> apt_buildings2.1 # bin by 10

apt_buildings %>%
  filter(!is.na(`non-smoking_building`)) %>%
  filter(!no_of_units %in% outliers) %>%
  mutate(year = as.Date(as.character(year_built), format="%Y")) %>%
  filter(year > as.Date("1910", format="%Y")) %>%
  mutate(year_category=cut(year_built, breaks=c(-Inf, seq(1905, 2020, 5)), labels=paste(seq(1900, 2015, 5)))) %>%
  group_by(year_category, property_type) %>%
  count(`non-smoking_building`) -> apt_buildings2.2 # bin by 5


apt_buildings %>%
  filter(!is.na(`non-smoking_building`)) %>%
  filter(!no_of_units %in% outliers) %>%
  mutate(year = as.Date(as.character(year_built), format="%Y")) %>%
  filter(year > as.Date("1910", format="%Y")) %>%
  group_by(year_built, property_type) %>%
  count(`non-smoking_building`) -> apt_buildings2.3 # bin by 1

apt_buildings2.1
apt_buildings2.2
apt_buildings2.3
```



#### 1.2.2.2 Task 8: Choose Bins

The task is to create 3 histograms out of summarized variables, with each histogram having different sized bins. Pick the "best" one and explain why it is the best. The binning for the first 2 plots were made manually due to the fact that `ggplot` is unable to make bins by year when in the `stat="identity"` mode. I had to use the `stat="identity"` mode since I am not actually using the counts of the data but rather taking the sum of *no_of_units* column by groups. Hence, the graphs are actually "self" made histograms made by `geom_bar`. Regardless, the goal which is to find which binning is appropriate can still be achieved. Looking at the 3 different bins, I believe that the "best" one would be the first one because the information is the same but with less groupings. However, it is still somewhat difficult to see if there is an apparent trend in the number of smoking buildings over the years.

```{r}
# task 8
# bin by 10
apt_buildings2.1 %>%
  ggplot(aes(x=year_category, y=n, fill=`non-smoking_building`)) +  
  geom_bar(stat="identity")

# bin by 5
apt_buildings2.2 %>%
  ggplot(aes(x=year_category, y=n, fill=`non-smoking_building`)) +  
  geom_bar(stat="identity")

# bin by 1
apt_buildings2.3 %>%
  ggplot(aes(x=year_built, y=n, fill=`non-smoking_building`)) +  
  geom_bar(stat="identity")
```

### 1.2.3 Research Question 3

The third research question is: Does the number of units change according to property type for certain visitor parking types?

#### 1.2.3.1 Task 2: Compute Observations

The task is to compute the number of observations for at least one of your categorical variables. Do not use the function `table()`! Similar to 1.2.2.1, it is more useful to use to actually sum over the *no_of_units* as it more accurately represents the apartment units buils over the groupings. I also calculate the *no_of_units_sum_log* as I wanted to see if the values work better on a log scale.

```{r}
# task 2
apt_buildings %>%
  filter(!is.na(visitor_parking)) %>%
  filter(!is.na(property_type)) %>%
  filter(!visitor_parking == "UNAVAILABLE") %>% 
  group_by(property_type, visitor_parking) %>%
  summarise(no_of_units_sum=sum(no_of_units)) %>%
  mutate(no_of_units_sum_log = log(no_of_units_sum))-> apt_buildings3
apt_buildings3
```

#### 1.2.3.2 Task 5: Graph Summary Statistics

The task is to create a graph out of summarized variables that has at least two geom layers. I created 2 graphs that plotted the *no_of_units_sum* by group where the x-axis represents the property type and the different lines represent parking type. If a certain property type does not affect the parking type, then the lines should be coincidental. It is clear that there is an effect because the lines are not coincidental. In addition, I plotted the log-scale of the sum of the number of units. It is even more evident that there is an effect.

```{r}
# task 5
apt_buildings3 %>% 
  ggplot(aes(x=property_type, y=no_of_units_sum, colour=visitor_parking, group=visitor_parking)) +
  geom_line() +
  geom_point()

apt_buildings3 %>% 
  ggplot(aes(x=property_type, y=no_of_units_sum_log, colour=visitor_parking, group=visitor_parking)) +
  geom_line() +
  geom_point()
```

### 1.2.4 Research Question 4

The fourth research question is: for apartments with indoor pools, does the total number of units change according to ward?

#### 1.2.4.1 Task 2: Compute Observations

The task is to compute the number of observations for at least one of your categorical variables. Do not use the function `table()`! I wanted to broaden the scope of the research question to counting the number of amenities by ward instead of just looking at the number of units that have pools by ward. I think it is more interesting to look at the overall number of amenities in each apartment building grouped by ward. In addition, I also grouped it by smoking vs non-smoking units because I think it would be interesting to see if there is a difference between the two.  Similar to the previous research questions, I compute the sum of the *no_of_amenities* over each group instead of just adding the counts of the number of observations since it is a more useful number to work with and it reflects the actual target characteristic. It is good to note that the number of amenities was first taken by processing the *amenities* column which is reflected in the first few lines of code below.

```{r}
# task 2
# count number of amenities by splitting words
no_of_amenities <- rep(0, nrow(apt_buildings))
for(i in 1:nrow(apt_buildings)) {
  amenities = apt_buildings$amenities[i]
  if(!is.na(amenities)) {
    no_of_amenities[i] = length(str_split(amenities, " ,")[[1]])
  }
}

# bind the number of amenities as a new column in the apt_buildings tibble df
cbind(apt_buildings, no_of_amenities) -> apt_buildings

apt_buildings %>%
  na.omit() %>%
  group_by(ward, `non-smoking_building`) %>%
  summarise(no_of_amenities_sum=sum(no_of_amenities)) -> apt_buildings4
apt_buildings4
```

#### 1.2.4.2 Task 5: Graph Summary Statistics

The task is to create a graph out of summarized variables that has at least two geom layers. I chose to highlight difference between the number of amenities present in the wards by smoking and non-smoking groups by creating 2 side-by-side graphs. This allows us to more easily see the differences in the shape and distribution. It is somewhat obvious that there is a difference.

```{r}
# task 5
apt_buildings4  %>%
  ggplot(aes(x=ward, y=no_of_amenities_sum, fill=`non-smoking_building`)) +
  geom_bar(stat="identity") +
  facet_wrap(~`non-smoking_building`)
```

## 1.3 Research Question Reflection

Here is the list of my research questions for quick reference:

1. Do certain property types prefer building apartments with a certain number of units?
2. Is there are trend on the number of units built in smoking apartment buildings from 1910 to 2020?
3. Does the number of units change according to property type for certain visitor parking types?
4. For apartments with indoor pools, does the total number of units change according to ward?

Below are my reflections so far.

**Based on the operations that you've completed, how much closer are you to answering your research questions?**


Based on the tasks completed, all of the research questions have been explored--however, some questions have shown to be more interesting than others. I think for all of them, the visualizations provide a very good overview of the general answers. However, it is hard to say for certain that we are confident in the "feel" that these visualizations give. It would be more helpful if some sort of hypothesis testing or modelling can be done. This is where the data can help.

**Think about what aspects of your research questions remain unclear.**

For research question (2), although I have decided which bin I would like to use to explore the trend, it is still unclear if there is a definite trend in the number of non-smoking versus smoking buildings that were built from 1910 to 2020. I think some sort of time series analysis can be done. It would also be good if I could potentially group by certain categorical variables that can affect this number. More data manipulations can be done to explore this topic. For research question (4), I decided to broaden the scope to counting the number of amenities grouped by ward and smoking type. Although I think this is interesting, I think it would be good if I could explore potential other variables that can affect this number. I think it would clarify why there could be a reason why this number varies over the wards.

    
**Can your research questions be refined, now that you've investigated your data a bit more?**

I think overall, the research questions have a good scope aside from the changes aspects that can be clarified as pointed out in point 2. I definitely think that as more data exploration is done, the research questions will definitely be more clear and the goal more defined. 


**Which research questions are yielding interesting results?**

Like I have mentioned, some research questions are more interesting than others. For example, I think research question (1) simply just asks if certain property types prefer building apartments with a certain number of units. I think it would be more interesting if the research question could broaden its scope to asking if certain property types prefer to build apartments with certain aspects like: window pane type or amenity type. Hence, I think it is more interesting to explore topics as mentioned in research question (2) or (4) since they already start looking at other variables that can affect the response variable. 


# 2 Tidy your data


A reminder of the definition of *tidy* data:

- Each row is an **observation**
- Each column is a **variable**
- Each cell is a **value**

*Tidy'ing* data is sometimes necessary because it can simplify computation. Other times it can be nice to organize data so that it can be easier to understand when read manually. 

## 2.1 Identifying Tidy data

First we take a look at the data using `glimpse`

```{r}
glimpse(apt_buildings)
```

There is clearly more than 8 columns, so we will only choose the columns that I mostly used for the research questions. I will store the data in the `apt_buildings5` data frame.

```{r}
apt_buildings %>%
  select(id,
         no_of_amenities,
         property_type,
         visitor_parking,
         ward,
         year_built,
         `non-smoking_building`,
         no_of_units,
         pets_allowed # added to meet the 8 minimum requirement
         ) %>%
  na.omit() -> apt_buildings5
head(apt_buildings5)
```

Now we answer if the data is Tidy by addressing the following points:

- Is each row an **observation**?
    
    It is clear that each row represents an observation where each observation is a specific apartment unit.
- Is each column a **variable**?
    
    It is also clear that each column represents a certain aspect of each observation.
- Is each cell a **value**?
    
    It can be seen that each cell represents a value taken on by each observation for that specific column variable. 

Hence, the data set `apt_buildings5` is tidy!

## 2.2 Tidy to untidy or vice versa

For reference, here is a view of the data set `apt_buildings5` as a Tidy data set.

```{r}
head(apt_buildings5)
``` 

I take the categorical variable *property_type* which take on the values PRIVATE, TCHC, and SOCIAL HOUSING and use each value as its own column variable which take on the value of the ward number. This means that if an apartment is has a property type that is PRIVATE and is found in ward 17, only the PRIVATE column will be filled while the other columns TCHC and SOCIAL HOUSING are empty. In addition, the PRIVATE cell for this apartment will be filled with the number 17. I store this untidy data set in a data frame called `apt_buildings6`. Below is my untidy operation along with an after look at the data frame. 


```{r}
apt_buildings5 %>%
  pivot_wider(names_from=property_type,
              values_from=ward) %>%
  as.data.frame()-> apt_buildings6
class(apt_buildings6)
head(apt_buildings6)
```

I choose to untidy in in this way because it would be interesting to see explore the grouping of property type by ward with respect to the continuous variables like *no_of_units* or *no_of_amenities*. Having it untidy in this way allows me to explore this grouping easier.

## 2.3 Choose Research Question



**Research Questions**
Based on my exploration and reflections, I choose the following research questions:
1. Research question 2: Is there are trend on the number of units built in smoking apartment buildings from 1910 to 2020? 
    
    Particularly, I think I can refine this question by asking what other factors can affect the trend in the number of smoking units versus non-smoking units over the years. I like this questions because I can perhaps create a time series model or even create more visualizations that include other factors in the analysis.
    
2. Research question 4: For apartments with indoor pools, does the total number of units change according to ward?

    As mentioned in 1.2.4.1, I have already broadened the scope of this question by changing it to counting the number of amenities in each ward. This is really an interesting research question because we could potentially factors that affect why certain apartment buildings can have gyms or swimming pools over other apartment buildings.
    
**Data set**

These 2 research questions are grouped differently from each other. The first one is grouped by years (binned by 10 years) and the second one is grouped by ward. It would make sense to keep the data as raw as possible ie. not much grouping or summarising already done before hand. Hence, I only selected the columns that interested me using the `select` function, got rid of unwanted outliers using `filter`, widened the non-smoking building variable with values from *no_of_units* using `pivot_wider` and changed the *year_built* variable from a string to a Date format using `mutate`. 

```{r}
apt_buildings %>%
  select(id,
         no_of_amenities,
         property_type,
         visitor_parking,
         ward,
         year_built,
         `non-smoking_building`,
         no_of_units,
         pets_allowed,
         heating_type,
         balconies,
         air_conditioning
         ) %>%
  filter(!no_of_units %in% outliers) %>%
  pivot_wider(names_from=`non-smoking_building`,
              values_from=no_of_units) %>%
  mutate(year = as.Date(as.character(year_built), format="%Y"))
```

